---
title: "ids_for_seq"
output: html_document
date: "2024-05-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/ahewett1/Documents')
library(tidyverse)
```

For individuals that died between hatching and fleding there are three sources in the database:
- ids with a ringid and a known death date (easy)
- ids with a ringid and no known death date (medium)
(These ids that have 'gone missing' so its not confirmed that they have survived but its unlikely they have survived if they go missing before the fledging date) 
- ids with no ringid and no known death date (difficult)

```{r}
clutch=read.csv("Barn_owl_general/Data_base_dfs/BarnOwls_Legacy_20231010153920_Clutch.csv")%>%
  select(Id, Season, Nestbox, MaleRing, FemaleRing, CrossFostered)

bird=read.csv("Barn_owl_general/Data_base_dfs/BarnOwls_Legacy_20231010153920_Bird.csv")%>%
  select(-RingType, -Species, -RingAdministrator,-CH1903X,-CH1903Y,
         -CH1903Altitude,-RingedBy,-GeneticSex,-PhenotypeSex,-RFRingId,
         -RFRingNumber,-RFRingDate,-RFRingDateErrorMargin,-Remarks
)


mes=read.csv("Barn_owl_general/Data_base_dfs/BarnOwls_Legacy_20231010153920_BirdMeasurement.csv")#%>%
    #select(RingId, HatchDate, BornClutchId, HatchClutchId, RaisedClutchId, RingDate, SiteId)


sequenced_ids=read.table("sequioa/3.inputfile_for_sequoia.raw", header=T)%>%
  select(IID)%>%
  rename(RingId=IID)%>%
  add_column(seq_round_1="yes")


lab=read.csv("Barn_owl_general/Data_base_dfs/Lab_20240523105823_LabIndividualTissue.csv")%>%
  select(IndividualIdentifier,OtherIdentifiers, SampleType,LabIndividualTissueId,TissueIdentifier,LabIndividualId)%>%
  rename(RingId=IndividualIdentifier)

lab_indiv=read.csv("Barn_owl_general/Data_base_dfs/Lab_20240523105823_LabIndividual.csv")%>%
    select(LabIndividualId,IndividualIdentifier,OtherIdentifiers)


```

## Ids with known death date ##

```{r}

age_dead_exact=bird%>%
  filter(!is.na(HatchDate) & !is.na(DeathDate) & DeathDate != "") %>%
  mutate(
    hatch_date = as.Date(HatchDate, format = "%d/%m/%Y"),
    death_date = as.Date(DeathDate, format = "%d/%m/%Y"),
    age_at_death = death_date - hatch_date)


age_dead_exact$age_at_death <- as.Date(as.character(age_dead_exact$death_date))-
  as.Date(as.character(age_dead_exact$hatch_date)) ## 

head(age_dead_exact)

```
748 ids with exact known death dates in the clutch that died before 90 days
```{r}
died_b4_90=age_dead_exact%>%
    filter(age_at_death<=90)

```
255 of these have no RingId (so died before they were big enough to be ringed)
493 have Ringid
```{r}
length(which(died_b4_90$RingId==""))
length(which(died_b4_90$RingId!=""))

```

How many have we already sequenced??:
(only ringed ids)
17 / 748
```{r}
samples_seqed=died_b4_90%>%
  left_join(sequenced_ids)%>%
  filter(seq_round_1=="yes")
head(samples_seqed)

```

Of ids that have been ringed:
239 ids with good quality DNA samples (blood / blood cell)

```{r}
to_seq=died_b4_90%>%
  filter(RingId!="")%>%
  left_join(lab)%>%
  filter(!(SampleType%in%c("swab","Swab", "Feather", "",NA)))

n_distinct(to_seq$RingId)
```
Also 44 ids with a 'space' in lab info so idk what that means??
```{r}
to_seq_space=died_b4_90%>%
  left_join(lab)%>%
  filter((SampleType%in%c("")))

n_distinct(to_seq_space$RingId)

```


Ids with known death dates and no ring id (n=255)

```{r}

no_ring_dd=died_b4_90%>%
  filter(RingId == "", 
        hatch_date > as.Date("2010-01-01")) # filtering for recent years (removes ~ 50 ids)


head(no_ring_dd)

```

quickly check others in clutch
(reminder to check full clutch as sometimes only one id per clutch have been sequenced)
```{r}
# sequenced_ids%>%
#   filter(RingId=='M026282')

seqd_clutches=bird%>%
  right_join(sequenced_ids, by = 'RingId')%>%
  distinct(BornClutchId)%>%
  na.omit()%>%
  add_column(others_in_clutch_seq='yes')

complete_fam=no_ring_dd%>%
  inner_join(seqd_clutches, by = 'BornClutchId')

complete_fam_clutch_ids=complete_fam%>%
  distinct(BornClutchId)


```










Ids with a ringID but no known death date

```{r}
full_clutch=read.csv("Barn_owl_general/Data_base_dfs/BarnOwls_Legacy_20231010153920_Clutch.csv")%>%
    select(Id, Season, Nestbox, MaleRing, FemaleRing, CrossFostered, UnhatchedEggs, InfertileEggs,     DeadEmbryos,BornEggs,BornAliveEggs,BornDeadEggs,Fledglings)
```

```{r}

ids_w_ids <- bird %>%
  select(RingId, BornClutchId) %>%
  filter(RingId != "") %>%
  group_by(BornClutchId) %>%
  summarise(num_w_ids_in_clutch = n(), .groups = 'drop')

num_fledlings <- full_clutch %>%
  filter(Season >2010) %>% ## 
  filter(CrossFostered != 1) %>% ## only for ids that were'nt cross fostered (otherwise )
  rename(BornClutchId = Id) %>%
  group_by(BornClutchId) %>%
  filter(Fledglings < BornAliveEggs) %>%
  right_join(bird, by = "BornClutchId") %>%
  select(RingId, BornClutchId, Fledglings, BornAliveEggs, BornEggs,HatchDate, DeathDate) %>%
  left_join(ids_w_ids, by = "BornClutchId") %>%
  filter(num_w_ids_in_clutch > Fledglings, !is.na(num_w_ids_in_clutch))

head(num_fledlings)

check_numbers=num_fledlings%>%
  select(BornClutchId, Fledglings, num_w_ids_in_clutch)%>%
  distinct(BornClutchId, .keep_all = T)%>%
  mutate(potential_number_new_ids=num_w_ids_in_clutch-Fledglings)

sum(check_numbers$potential_number_new_ids)

```

```{r}
num_fledlings[1:4,1:8]
```

```{r}
full_clutch%>%
  select(Id, Fledglings, BornAliveEggs, BornEggs)%>%
  arrange(Id)%>%
  filter(Id>0)
```





Ids that died in clutch but still may have lab stuff 


Clutches where the number of born eggs doesnt equal the number of of born alive eggs 
i.e. some eggs have died for one reason or another 
```{r}
eggs_unhatched=full_clutch%>%
  rename(BornClutchId=Id)
  
```

```{r}
comb_bird=bird%>%
  left_join(eggs_unhatched)%>%
    filter(BornEggs!=BornAliveEggs)
```
## Birds with no Ringid (so died before they were big enough to be ringed) ##

```{r}
no_ring=bird=read.csv("Barn_owl_general/Data_base_dfs/BarnOwls_Legacy_20231010153920_Bird.csv")%>%
  filter(RingId=="")

head(no_ring)

```

