---
title: "pedigree_diagnostics"
output:
  pdf_document: default
  html_document: default
date: "2024-02-22"
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/ahewett1/Documents")

library(tidyverse)
library(sequoia)

```

# Pedigree reconstruction using Sequoia 

Pedigree reconstruction was done in two stages - as recommended in the sequoia vignette.

In both stages reconstructed pedigrees were compared to the corrected pedigree (i.e. the one already purged by Alex and Elu, so hopefully shouldn't be too many mismatches initially!)

1. *Parentage assignment* 
 + Assigns genotyped parents to genotyped offspring
 + Very quick, designed to find some initial errors in the field pedigree
2. *Full reconstruction* 
  + Creates dummy ids where parent doesnt match offspring and cannot find a match in the database ... also useful to find missing links
  + Multiple iterations to generate a ML of parentage 


# Summary #

- Initial parentage assignment showed some ids where genetic sex and recorded sex didnt match i.e. so mother was actually the father 
- Some examples of mislabeled ids
- 1 possible extra pair paternity 
- 1 id where matched to completely different parents from a few years ago. Could be a pair of ids that were sampled and ringed then left the study area and tags fell off, then re-entered the study area 5 yrs later and were ringed again.(?) 
- A few examples of mothers/fathers being mismatched with offspring and have a relatedness O-P of ~0.25 i.e. more like auntie-niece etc. Checked the mothers and many of them come from crossfostered nests, so perhaps a mix up in ids i.e. where sibling was genotyped under the wrong id. But not sure when sampling Vs ringing occurs. 

.... Details below

(When i use the terms recorded and hypothesised im referring to the parents recorded in the pedigree vs the parents inferred by sequoia)


# Inital parentage assignment

Only 20 mismatches between the reconstructed pedigree and the original pedigree

```{r, echo=FALSE}
load("sequioa/pedigree_FULL_recontr_out26.2.RData")

ped_gen_cor=read.table("sequioa/pedigreeCORRECTED.tab", header = T)
parentage_ped=seq_parentage_out$Pedigree


sequenced_ids_read=read.table("sequioa/3.inputfile_for_sequoia.raw", header=T)%>%
  select(IID)
sequenced_ids=as.character(sequenced_ids_read[,1])


compare_parentage=PedCompare(
  Ped1 = ped_gen_cor,
  Ped2 = parentage_ped,
  DumPrefix = c("XF", "XM"),
  SNPd = sequenced_ids
  
)

```

### Most initial mismatches appear to be a mix up in sex or mislabelling 
- M026267 and M038112  recorded as female but actually a male according to the genetic sex
- M026658 recorded as male but genetic sex is female
- A potential mislabelling of id 'M031195' which is actually 'M041195'??

**Below shows the flagged mismatches from the comparison:**

- columns dam.1 and sire.1 are the recorded mother and father 
- columns dam.2 and sire.2 are the inferred parents from seqouia
```{r}

mismatch_parents=compare_parentage$Mismatch
print(mismatch_parents)
```
## Changed these ids and re-run parentage check ##
```{r, echo=FALSE}
ped_corrected=read.table("sequioa/pedigreeCORRECTED.tab", header = T)%>%
  mutate(dadid=case_when(
    momid=="M026267" ~"M026267", 
    momid=="M038112" ~ "M038112",
    dadid=="M026658" ~ NA,
    dadid=="M031195" ~ "M041195",
    TRUE ~ dadid
  ))%>%
    mutate(momid=case_when( ## found during first parentage assigment
      dadid=="M026658" ~ "M026658",## M026267 , M038112  recorded as female but actually a male 
      momid=="M026267" ~ NA,       ## M026658 recorded as male but genetic sex is female
      momid=="M038112" ~ NA,
      TRUE ~ momid 
      
    ))
```

## Only 7 parentage mismatches remaining ##

- 7 mismatches removed due to wrong sex
- 3 mismatches removed due to mislabelled id 
- 4 of the 7 remaining mismatches comes from the same mother.  
- 2 of the 7 remaining mismatches are in the same id

```{r, echo=FALSE}

compare_parentage=PedCompare(
  Ped1 = ped_corrected,
  Ped2 = parentage_ped,
  DumPrefix = c("XF", "XM"),
  SNPd = sequenced_ids
  
)

remaining_mismatch_parents=compare_parentage$Mismatch
print(remaining_mismatch_parents)
```
```{r, echo=FALSE}
GRM=readRDS("sequioa/All3085_AUTOSAUMES_RP502SNPs.RDS")
```

## Double parent mismatch for id M011867 ##

#### Comparing parental mismatches with GRM ####


- Recorded parents **are not genotyped** (M011634, M011635)

- High genomic relatedness (~0.5) with suggested parents (871649 and 870404) 

```{r}
GRM['M011867', '871649']
GRM['M011867', '870404']

```

### Comparing with database info: ###

```{r, echo=FALSE}
df_bird=read.csv("Barn_owl_general/BarnOwls_Legacy_20231010153920/BarnOwls_Legacy_20231010153920_Clutch.csv")%>%
  select(Id, Season, Nestbox, MaleRing, FemaleRing, CrossFostered)
```

Recorded parents:
```{r, echo=FALSE}
df_bird%>%
  filter(FemaleRing=="M011634") #recorded mother
```

Hypothesised parents
```{r,echo=FALSE}
df_bird%>%
  filter(FemaleRing=="871649") # hypothesised mother
```

```{r,echo=FALSE}
life2=read.csv("Barn_owl_general/BarnOwls_Legacy_20231010153920/BarnOwls_Legacy_20231010153920_Bird.csv")%>%
    select(RingId, HatchDate, BornClutchId, HatchClutchId, RaisedClutchId, RingDate, SiteId)


```

According to the hatch date and clutch id its clear that the recorded mother and father ID is correct as it matches the clutches they have

ID (M011867) was born in 2008, ~5 years after the hypothesised parents had a clutch. 
```{r, echo=FALSE}
life2%>%
  filter(RingId=="M011867")
```

Could it be possible that father M011635 = 870404 and mother M011634 = 871649 but came back into the study area 5 years later??
That would mean breeding for >10 years?

M011867 has 3 siblings, none of which have been genotyped (checked but not shown)
```{r, echo=FALSE}
life2%>%
  filter(BornClutchId=="1027")
```

#### >> can we check the photos to see if they look like the same individuals?? ####


## M040592	 ##
- Has a mismatched sire 

Hypothesised sire has relatedness of 0.467 with offspring, recorded sire not genotyped

```{r, eval==FALSE}

print(remaining_mismatch_parents)

#recorded sire 
#GRM['M040592', 'M043668'] ## 

```

```{r}
#hypothesised sire
GRM['M040592', 'M040587']

```

Clutches where recorded sire is the father in the database:
```{r, echo=FALSE}
df_bird%>%
  filter(MaleRing=="M043668") #recorded sire
```

Clutches where hypothesised sire is the father in the database:
```{r, echo=FALSE}
df_bird%>%
  filter(MaleRing=="M040587") # hypothesised sire
```

Id clutch:
Born in 2019 so actually overlaps both male seasons
```{r, echo=FALSE}
life2%>%
  filter(RingId=="M040592")
```

Other ids in clutch 2005:
```{r}
life2%>%
  filter(BornClutchId=="2005")

```

Relatedness between hypothesised father and another offsrping in the same clutch
```{r}
GRM['M040590', 'M040587']
```

A sibling in the clutch has been genotyped as well and also has a very high relatedness with the hypothesised father (0.4795)
So potentially the whole clutch has the wrong recorded father. 
(strange that the sibling didnt come back with an error though, maybe paternal error rates are too high?)

So, clear that the recorded father raised the clutch 
Could be extra-pair paternity as they occur in very similar season? so M028977 mated with M040587 but M043668 did the raising

#### >> Can we check how far away the 2 nestboxes are: 2 and 167B ####


```{r, echo=FALSE}
print(remaining_mismatch_parents)

```

## Dam M027000 ##
- Has many mismatches with her recorded offspring
- But has not been genotyped 
- Hypothesised mother (M022696) has high genomic relatedness with all clutch offspring (this is why it was flagged as a mismatch)

```{r}
GRM['M026190', 'M022696']
GRM['M026231', 'M022696']
GRM['M026232', 'M022696']
GRM['M010989', 'M022696']

```

Recorded mother clutch info:

Has had multiple seasons of offspring
```{r, echo=FALSE}
df_bird%>%
  filter(FemaleRing=="M027000") #recorded 

```
Hypothesised mother clutch info:

Has not had any offspring according to database
```{r, echo=FALSE}
df_bird%>%
  filter(FemaleRing=="M022696") # hypothesised 
```

Origin clutch of recorded mother:
```{r, echo=FALSE}
life2%>%
  filter(RingId=="M027000")
```
Origin clutch of hypothesised mother:
```{r, echo=FALSE}

life2%>%
  filter(RingId=="M022696")
```

- Recorded mother and inferred mother are siblings and from the same clutch and moved to the same clutch 
- Possibly some sort of mix up between the ids after cross fostering??
- or very high relatedness between auntie and all her nephews/nieces which is kind of unlikely?



# Full reconstruction #

```{r, echo=FALSE}
full_ped_recontr=seq_reconstr_out$Pedigree

full_compare_parentage=PedCompare(
  Ped1 = ped_corrected,
  Ped2 = full_ped_recontr,
  DumPrefix = c("XF", "XM"),
  SNPd = sequenced_ids
  
)

full_remaining_mismatch_parents=full_compare_parentage$Mismatch
##print(full_remaining_mismatch_parents)

```
```{r, echo=FALSE}
##print(remaining_mismatch_parents)
seq_reconstr_out$DummyIDs

```

- Removed IDs already discussed

```{r, echo=FALSE}
full_remaining_mismatch_parents_filt=full_remaining_mismatch_parents%>%
  filter(sire.class=="Mismatch"|dam.class=="Mismatch")%>%
  filter(dam.1!="M027000")%>%
  filter(sire.1!="M043668")%>%
  filter(dam.1!="M011634")

```

- 36 other examples where parents are mismatched 

```{r, echo=FALSE}
##full_remaining_mismatch_parents_filt
```
## Start just with mismatched sires ##

All ids where the recorded sire is mismatched with offspring i.e. parental genotype doesnt work with offspring genotype

Have all had dummy id's created as the sires (hence mismatched)

```{r, echo=FALSE}
full_remaining_mismatch_parents_filt%>%
  filter(sire.class=="Mismatch")%>%
  filter(id.sire.cat=="GG")
```

## Relatedness between recorded sire and offspring ##

- Many as expected for parent-offspring so probably incorrect by sequoia / genotyping coverage 
```{r}
GRM['M005013', '899009']
```
```{r}
GRM['M022861', 'M006413']
GRM['M022863', 'M006413']
GRM['M022864', 'M006413']

GRM['M026547', 'M006413']
GRM['M026790', 'M006413']
GRM['M026792', 'M006413']
GRM['M026793', 'M006413']
GRM['M026848', 'M006413']

GRM['M031661', 'M022855']
GRM['M032001', 'M006413']
GRM['M038183', 'M031882']

```

## Only 2 flagged mismatches show quite low relatedness between parent-offspring ##
- M032425
- M040751

## M032425 ##

Actually M032425 has quite low relatedness with both parents which is weird:
(but not super low )

```{r}
GRM['M032425', 'M032203'] 
GRM['M032425', 'M032228'] 

```

Details of focal id clutch
```{r, echo=FALSE}
life2%>%
  filter(RingId=="M032425")

```
Details of sire clutch:
```{r, echo=FALSE}
life2%>%
  filter(RingId=="M032203")
```

Details of mother clutch:
(note this wasnt flagged as a mismatch but relatedness was very low so wanted to take a look)
```{r, echo=FALSE}
life2%>%
  filter(RingId=="M032228")
```

```{r, eval=FALSE}
df_bird%>%
  filter(Id=="1463")

df_bird%>%
  filter(Id=="1465")
```

## M040751 ##
Relatedness low for parent-offspring
```{r}
GRM['M040751', 'M040585']

```
Clutch info for focal id
```{r, echo=FALSE}

life2%>%
  filter(RingId=="M040751")
```

Clutch info for recorded sire:
(no info on born clutch etc so immigrant id)
```{r, echo=FALSE}
life2%>%
  filter(RingId=="M040585")
```

Double checking nextboxID 
```{r, echo=FALSE}
df_bird%>%
  filter(Id=="2094")

```
#### All records look to be correct in the database, so not sure whats going on here! ####

## Mismatched mothers ##
```{r, echo=FALSE}
full_remaining_mismatch_parents_filt%>%
  filter(dam.r=="nomatch")
```
All these seem fine (again maybe an error with sequoia/genotyping)
```{r}
GRM['M005026', '877475']
GRM['M005473', '877475']

GRM['M026986', 'M022177']
GRM['M032144', 'M027970']
GRM['M032334', 'M026986']

```

## These are too low for parental relatedness ##
```{r}

GRM['M026444', 'M026258']

GRM['M026522', 'M026392']
GRM['M026523', 'M026392']
GRM['M026524', 'M026392']
GRM['M026525', 'M026392']

```

```{r, echo=FALSE}
# life2%>%
#   filter(RingId=="M026444")
```

## M026444 ##
Recorded mother clutch origin info:
```{r}
life2%>%
  filter(RingId=="M026258")
```

```{r, echo=FALSE}
# df_bird%>%
#   filter(Id=="1355")
```

Mother was raised in a different clutch than she was born, and was ringed almost a month after hatching. 
Could the individuals have got confused somehow? 
so was actually her sibling that was sampled? (that would make sense with relatedness)


```{r, echo=FALSE}
# life2%>%
#   filter(RingId=="M026522")
# 
# life2%>%
#   filter(RingId=="M026523")
# 
# life2%>%
#   filter(RingId=="M026524")
# 
# life2%>%
#   filter(RingId=="M026525")
# 
# 
# df_bird%>%
#   filter(Id=="1308") #hatched clutch
# 
# df_bird%>%
#   filter(Id=="1315") #raised clutch

```
Recorded mother origin:
```{r, echo=FALSE}
## recorded mother
life2%>%
  filter(RingId=="M026392")
```
Checking mothers origin clutch
```{r, echo=FALSE}
##check mothers origin clutch
life2%>%
  filter(BornClutchId=="1224")

```
Again mother has been moved, so could she have got mixed up with a sibling??
Mother is ringed >1 month after hatching ... is this before or after theyve been moved and sampled etc.?
And why is the ring date so different between the siblings in this clutch?

## Checking hypothesis about sibling mismatches ##
```{r, echo=FALSE}
mes=read.csv("Barn_owl_general/BarnOwls_Legacy_20231010153920/BarnOwls_Legacy_20231010153920_BirdMeasurement.csv")#%>%
    ##select(RingId, HatchDate, BornClutchId, HatchClutchId, RaisedClutchId, RingDate, SiteId)
```

#### ID M026392 ####
example from earlier

```{r, echo=FALSE}
life2%>%
  filter(RingId=="M026392")
```
Ringed on 22/09/2011

But has measurements starting at 12/08/2011
Does this mean it was sampled on that date? In which case could the ids have been mixed up??

```{r, echo=FALSE}
mes%>%
  filter(RingId=="M026392")%>%
  head(n = c(7, 8))
```


### M032203 ###
Another example from above

Same story. Has measurements taken on 2/07/2013 but was ringed on 15/07/2013. 
So if it was sampled at the same time the first measurement was taken then could have been mixed up with the sibling, which would explain the relatedness of ~0.25 between recorded parent and offspring. 

```{r, echo=FALSE}
life2%>%
  filter(RingId=="M032203")
```
```{r, echo=FALSE}
mes%>%
  filter(RingId=="M032203")%>%
    head(n = c(7, 8))

```

```{r, echo=FALSE}
life2%>%
  filter(RingId=="M032228")
```
```{r, echo=FALSE}
mes%>%
  filter(RingId=="M032228")%>%
    head(n = c(7, 8))

```

M032228
```{r}
GRM['M032425', 'M032228']
```

