---
title: "For_seq"
output: html_document
date: "2024-06-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/ahewett1/Documents')
library(tidyverse)
```

```{r, echo=FALSE}
clutch=read.csv("Barn_owl_general/Data_base_dfs/BarnOwls_Legacy_20231010153920_Clutch.csv")%>%
  select(Id, Season, Nestbox, MaleRing, FemaleRing, CrossFostered)

bird=read.csv("Barn_owl_general/Data_base_dfs/BarnOwls_Legacy_20231010153920_Bird.csv")%>%
  select(-RingType, -Species, -RingAdministrator,-CH1903X,-CH1903Y,
         -CH1903Altitude,-RingedBy,-GeneticSex,-PhenotypeSex,-RFRingId,
         -RFRingNumber,-RFRingDate,-RFRingDateErrorMargin,-Remarks
)
mes=read.csv("Barn_owl_general/Data_base_dfs/BarnOwls_Legacy_20231010153920_BirdMeasurement.csv")#%>%
    #select(RingId, HatchDate, BornClutchId, HatchClutchId, RaisedClutchId, RingDate, SiteId)

sequenced_ids=read.table("sequioa/3.inputfile_for_sequoia.raw", header=T)%>%
  select(IID)%>%
  rename(RingId=IID)%>%
  add_column(seq_round_1="yes")

lab=read.csv("Barn_owl_general/Data_base_dfs/Lab_20240523105823_LabIndividualTissue.csv")%>%
  select(IndividualIdentifier,OtherIdentifiers, SampleType,LabIndividualTissueId,
         TissueIdentifier,LabIndividualId, BoxCode, BoxName, TubeIdentifier)%>%
  rename(RingId=IndividualIdentifier)

lab_indiv=read.csv("Barn_owl_general/Data_base_dfs/Lab_20240523105823_LabIndividual.csv")%>%
    select(LabIndividualId,IndividualIdentifier,OtherIdentifiers)

big_boy_lab=read.csv("Barn_owl_general/Data_base_dfs/Ta_Switzerland_Tissue_DNA _BIG_FILE_11.01.24.csv")
```

Types of tissue samples:
(Will be prioritising blood and blood cell)
```{r}
table(lab$SampleType)
```


# ~~ 1.1 ~~ #
 
### Dead ids with: ###
- Known (or in some cases inferred) death date and birth date (i.e. those born in the study area) 
- A ring id 
- Born in or after 2010
```{r}
## getting individuals with known death and birth dates 
age_dead_exact=bird%>%
  filter(!is.na(HatchDate) & !is.na(DeathDate) & DeathDate != "") %>%
  mutate(
    hatch_date = as.Date(HatchDate, format = "%d/%m/%Y"), #mutating dates to correct formatting
    death_date = as.Date(DeathDate, format = "%d/%m/%Y"),
    age_at_death = death_date - hatch_date)

dead_wRing_knownDd_blood=age_dead_exact%>%
    filter(age_at_death<=90, #died before 90 days
           RingId != "", # has ring id
          hatch_date > as.Date("2010-01-01"))%>% # hatched after / in 2010
  anti_join(sequenced_ids, by = "RingId")%>% # remove ids sequenced last time
  left_join(lab,  by = "RingId")%>% #joing with lab info
  filter(SampleType%in%c("blood","Blood cell"))%>% #check which ones have blood samples
  distinct(RingId, .keep_all = TRUE) #

```
### 462 ids in total but of these, number of ids with blood samples: ###
```{r}
n_distinct(dead_wRing_knownDd_blood$RingId)
```

# ~~ 1.2 ~~ #

### Siblings of the ids that died ###
.. just to complete the family and add power to analysis
 
```{r}
# getting born clutch id of those in stage 1.1
clutches_of_deads1=dead_wRing_knownDd_blood%>%
  select(BornClutchId)%>% 
    distinct(BornClutchId)
```
Number of different clutches
```{r, message=FALSE}
n_distinct(clutches_of_deads1$BornClutchId)

all_sibs=bird%>% # all records of all ids 
  right_join(clutches_of_deads1)%>% # getting only ids born in the clutches of interest (645 ids) 
  anti_join(sequenced_ids, by = "RingId")%>% # removing ids already sequenced
  anti_join(dead_wRing_knownDd_blood, by = 'RingId') # removing ids from 1.1

ringed_sibs=
  all_sibs%>%
  left_join(lab)%>% # combing with lab info 
  filter(SampleType%in%c("blood","Blood cell")) # filtering for sibs only with blood samples.
```
### Number of ids with suitable blood: ###
```{r}
n_distinct(ringed_sibs$RingId)
```

# ~~ 1.3 ~~ # 

### Siblings of previously sequenced individuals that have high inbreeding coefficients ###

```{r, message=FALSE}
##Read in inbreeding coefficients
inb=read.table("Inbreeding_depression_owls/All3085_FuniWE_FHBD512g.txt", header = T)%>%
  rename(RingId=INDVs)
#top 90% inbred individuals from 1st round of sequencing
top_ibcs=inb%>%
  filter(FuniWE > quantile(FuniWE, 0.9))%>% # top % ibcs
  left_join(bird)%>% ## joining with bird info to get DOB
  filter(as.Date(HatchDate, format = "%d/%m/%Y") > as.Date("2010-01-01")) # hatched after / in 2010

n_distinct(top_ibcs$BornClutchId) ##  number of different clutches that have highest ibcs 

#inbred clutches
inbred_clutches=top_ibcs%>%
  select(BornClutchId)%>%
  distinct(BornClutchId)%>%
  na.omit()
# inbred sibling within those inbred clutches
inbred_sibs=bird%>%
  right_join(inbred_clutches)%>% #
  anti_join(sequenced_ids, by = "RingId")%>% # remove ids previously sequenced
  anti_join(dead_wRing_knownDd_blood, by = 'RingId')%>% # remove ids from stage 1.1
  anti_join(ringed_sibs, by = 'RingId')%>% # remove ids from stage 1.2
  left_join(lab,  by = "RingId")%>% # join with lab info
  filter(SampleType%in%c("blood","Blood cell")) #filter for only ids with blood samples 
```
### Number of ids with blood samples: ###

```{r}
n_distinct(inbred_sibs$RingId) # number of samples 
```

# Total for part 1 #

## Including: ##

- Confirmed dead ids that died within 90 days of hatching (224)
-  ... Their siblings (89)
- Siblings of previously sequenced ids that had high inbreeding coefficients (84)

All of which are born after 2010 (as this is when regular visits started) and have good blood samples

Total:

```{r}
224 + 84 + 89
```
#############
## PLATE 2 ##
#############

## ~~ 2.1 ~~ ##

Individuals that went missing before 45 days old, when others in the clutch were measured after and the last time they were observed they were <35 days old. For non-cross fostered ids so far. 
These ids have also been manually checked on the database using timeline and 'possibly dead' box. 
Only for ids with blood. 
See cacl_missing.R for full run through,

```{r}
missings=read.csv("Barn_owl_general/missing_ids/Confirmed_missingIDs_with_blood_and_rings.csv")%>%
  anti_join(sequenced_ids, by = "RingId")%>% # remove ids previously sequenced (n=1)
  anti_join(dead_wRing_knownDd_blood, by = 'RingId')%>% # remove ids from stage 1.1 (n=0)
  anti_join(ringed_sibs, by = 'RingId') # remove ids from stage 1.2 (n=9)
  
  
n_distinct(missings$RingId)

```

## ~~ 2.2 ~~ ##

Siblings of missing ids 


```{r}

missings_clutches=missings%>%
  select(BornClutchId)%>%
  distinct(BornClutchId)%>%
  na.omit()
# inbred sibling within those inbred clutches
missing_sibs=bird%>%
  right_join(missings_clutches)%>% #
  anti_join(missings, by = "RingId")%>% # remove ids previously sequenced
  anti_join(sequenced_ids, by = "RingId")%>% # remove ids previously sequenced
  anti_join(dead_wRing_knownDd_blood, by = 'RingId')%>% # remove ids from stage 1.1
  anti_join(ringed_sibs, by = 'RingId')%>% # remove ids from stage 1.2
  left_join(lab,  by = "RingId")%>% # join with lab info
  filter(SampleType%in%c("blood","Blood cell")) 

n_distinct(missing_sibs$RingId)


```




## ~~ 2.3 ~~ ##


### All individuals born after 2010 with no RingId ###
(i.e. ids that died before they were old enough to be ringed)

```{r}
All_after2010_noRing=bird%>%
    filter(as.Date(HatchDate, format = "%d/%m/%Y") > as.Date("2010-01-01"))%>% # hatched after / in 2010
  filter(RingId=="")
 
n_distinct(All_after2010_noRing$BirdId)


```

Of those, those with a definite known death date:
(others probably went missing so dont have a definite death date)

```{r}

clutch_born=clutch%>%
    rename(BornClutchId=Id)

clutch_raised=clutch%>%
    rename(RaisedClutchId=Id)

dead_noRing_knownDd=All_after2010_noRing%>%
  filter(!(DeathDate==""))%>%
  left_join(clutch_born, by = 'BornClutchId')%>%
  select(BirdId, Mark, BornClutchId, Season, Nestbox.y, CrossFostered, HatchDate, DeathDate)

n_distinct(dead_noRing_knownDd$BirdId)

```

And attempt to link to sample info 

```{r}

NBs=big_boy_lab%>%
  filter(str_starts(Sample_Name_1, 'NB')| str_starts(Sample_Name_2, 'NB'))%>%
  filter(!str_starts(Sample_Name_2, 'M') &  !str_starts(Sample_Name_3, 'M') & !str_starts(Sample_Name_1, 'M'))%>%
  filter(Tissue.Type%in%c("blood","Blood"))#check which ones have blood samples
```

First sample same seq: NB, Season, N, nestbox, Mark
and match to sample name 2

```{r}
dead_noRing_knownDd_NB_S_N_nest_M=dead_noRing_knownDd%>%
 # filter(CrossFostered==0 | is.na(CrossFostered))%>%
  drop_na(BornClutchId)%>% #remove 2 ids with no clutch info
  mutate(nestbox_edit=gsub("[A-Za-z]", "", Nestbox.y))%>%
  mutate(Sample_Name_2 = paste0('NB', Season, 'N', nestbox_edit, Mark))


sample_name2_match_NB_S_N_nest_M=dead_noRing_knownDd_NB_S_N_nest_M%>%
  inner_join(NBs, by = 'Sample_Name_2')

n_distinct(sample_name2_match_NB_S_N_nest_M$BirdId)

```



NB, nestbox, Mark
and match to sample name 1 & 2

```{r}
dead_noRing_knownDd_NB_nest_mark=dead_noRing_knownDd%>%
  drop_na(BornClutchId)%>% #remove 2 ids with no clutch info
  mutate(nestbox_edit=gsub("[A-Za-z]", "", Nestbox.y))%>%
  mutate(Sample_Name_2 = paste0('NB', nestbox_edit,Mark))

sample_name2_match_NB_nest_mark=dead_noRing_knownDd_NB_nest_mark%>%
  inner_join(NBs, by = 'Sample_Name_2',relationship = "many-to-many")

n_distinct(sample_name2_match_NB_nest_mark$BirdId)

## sample name 1

dead_noRing_knownDd_NB_nest_mark=dead_noRing_knownDd%>%
  drop_na(BornClutchId)%>% #remove 2 ids with no clutch info
  mutate(nestbox_edit=gsub("[A-Za-z]", "", Nestbox.y))%>%
  mutate(Sample_Name_1 = paste0('NB', nestbox_edit,Mark))

sample_name1_match_NB_nest_mark=dead_noRing_knownDd_NB_nest_mark%>%
  inner_join(NBs, by = 'Sample_Name_1',relationship = "many-to-many")

n_distinct(sample_name1_match_NB_nest_mark$BirdId)

```

N, nestbox, Mark
and match to sample name 1


```{r}

dead_noRing_knownDd_N_nest_M=dead_noRing_knownDd%>%
 # filter(CrossFostered==0 | is.na(CrossFostered))%>%
  drop_na(BornClutchId)%>% #remove 2 ids with no clutch info
  mutate(nestbox_edit=gsub("[A-Za-z]", "", Nestbox.y))%>%
  mutate(Sample_Name_1 = paste0('N', nestbox_edit, Mark))




sample_name1_match_N_nest_M=dead_noRing_knownDd_N_nest_M%>%
  inner_join(NBs, by = 'Sample_Name_1',relationship = "many-to-many")

n_distinct(sample_name1_match_N_nest_M$BirdId)
```




```{r}
all_unringed_linked_blood=rbind(sample_name2_match_NB_S_N_nest_M,sample_name1_match_NB_nest_mark)%>%
  rbind(sample_name2_match_NB_nest_mark)%>%
  rbind(sample_name1_match_N_nest_M)

n_distinct(all_unringed_linked_blood$BirdId)


```







## ~~ 2.4~~ ##




### Parents of clutches in 1.1 and 1.3 ###

```{r}
## parents from 1.1 and 1.2
parents_of_dead1s=clutches_of_deads1%>%
  rename(Id=BornClutchId)%>%
  left_join(clutch, by = 'Id') %>% ## joining with clutch info
  pivot_longer(cols = c(MaleRing, FemaleRing), values_to = "RingId") #putting parents into one col


n_distinct(parents_of_dead1s$RingId) # number of samples 


parents_of_inbred1s=inbred_clutches%>%
  rename(Id=BornClutchId)%>%
  left_join(clutch, by = 'Id')%>%
  pivot_longer(cols = c(MaleRing, FemaleRing), values_to = "RingId")%>%#putting parents into one col
  filter(CrossFostered==0|is.na(CrossFostered)) ## also add na 


n_distinct(parents_of_inbred1s$RingId) # number of samples 

all_parents_inbred1s_and_dead1s = c(parents_of_dead1s$RingId, parents_of_inbred1s$RingId)

# dataframe with all inbred and dead ones parents fromm 1.1 and 1.2
all_parents_inbred1s_and_dead1s_df = data.frame(RingId = all_parents_inbred1s_and_dead1s)%>%
  filter(RingId!='')%>%
  distinct(RingId, .keep_all = TRUE) #

all_parents=all_parents_inbred1s_and_dead1s_df%>%
  anti_join(missings, by = "RingId")%>% # remove ids from 2.1
  anti_join(sequenced_ids, by = "RingId")%>% # remove ids previously sequenced
  anti_join(dead_wRing_knownDd_blood, by = 'RingId')%>% # remove ids from stage 1.1
  anti_join(ringed_sibs, by = 'RingId')%>% # remove ids from stage 1.2
  left_join(lab,  by = "RingId")%>% # join with lab info
  filter(SampleType%in%c("blood","Blood cell")) 

n_distinct(all_parents$RingId) # number of samples 

  
```















