---
title: "For_seq"
output: html_document
date: "2024-06-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/ahewett1/Documents')
library(tidyverse)
```

```{r}
clutch=read.csv("Barn_owl_general/Data_base_dfs/BarnOwls_Legacy_20231010153920_Clutch.csv")%>%
  select(Id, Season, Nestbox, MaleRing, FemaleRing, CrossFostered)

bird=read.csv("Barn_owl_general/Data_base_dfs/BarnOwls_Legacy_20231010153920_Bird.csv")%>%
  select(-RingType, -Species, -RingAdministrator,-CH1903X,-CH1903Y,
         -CH1903Altitude,-RingedBy,-GeneticSex,-PhenotypeSex,-RFRingId,
         -RFRingNumber,-RFRingDate,-RFRingDateErrorMargin,-Remarks
)
mes=read.csv("Barn_owl_general/Data_base_dfs/BarnOwls_Legacy_20231010153920_BirdMeasurement.csv")#%>%
    #select(RingId, HatchDate, BornClutchId, HatchClutchId, RaisedClutchId, RingDate, SiteId)

sequenced_ids=read.table("sequioa/3.inputfile_for_sequoia.raw", header=T)%>%
  select(IID)%>%
  rename(RingId=IID)%>%
  add_column(seq_round_1="yes")

lab=read.csv("Barn_owl_general/Data_base_dfs/Lab_20240523105823_LabIndividualTissue.csv")%>%
  select(IndividualIdentifier,OtherIdentifiers, SampleType,LabIndividualTissueId,
         TissueIdentifier,LabIndividualId, BoxCode, BoxName, TubeIdentifier)%>%
  rename(RingId=IndividualIdentifier)

lab_indiv=read.csv("Barn_owl_general/Data_base_dfs/Lab_20240523105823_LabIndividual.csv")%>%
    select(LabIndividualId,IndividualIdentifier,OtherIdentifiers)
```

Types of tissue samples:
(Will be prioritising blood and blood cell)
```{r}
table(lab$SampleType)
```



 ## STAGE 1 ##
 
Dead ids with known death date and a ring id born in or after 2010:
462 ids 

With blood or blood cell sample:
224 ids 

```{r}
age_dead_exact=bird%>%
  filter(!is.na(HatchDate) & !is.na(DeathDate) & DeathDate != "") %>%
  mutate(
    hatch_date = as.Date(HatchDate, format = "%d/%m/%Y"),
    death_date = as.Date(DeathDate, format = "%d/%m/%Y"),
    age_at_death = death_date - hatch_date)

## dead ids with known death date and known ring id (post 2010)
dead_wRing_knownDd_blood=age_dead_exact%>%
    filter(age_at_death<=90, #died before 90 days
           RingId != "", # has ring id
          hatch_date > as.Date("2010-01-01"))%>% # hatched after / in 2010
  left_join(sequenced_ids, by = "RingId")%>%
  filter(is.na(seq_round_1)) %>% ## werent sequeced last time 
  left_join(lab,  by = "RingId")%>%
  filter(SampleType%in%c("blood","Blood cell"))%>%
  distinct(RingId, .keep_all = TRUE) # remove repeated records of ids (where there are multiple blood samples)

n_distinct(dead_wRing_knownDd_blood$RingId)

LAB_info=dead_wRing_knownDd_blood%>%
  select(RingId, LabIndividualId, LabIndividualTissueId, SampleType, BoxCode, BoxName, TubeIdentifier)
  
  
dead_wRing_knownDd_blood_IDs=
  dead_wRing_knownDd_blood%>%
  select(RingId)%>%
  add_column(proposed_seq_round_1="yes")

head(dead_wRing_knownDd_blood_IDs)
```

 ## STAGE 1.2 ##
 Siblings of the ids that died (with blood)
 
 89 with suitable blood 

```{r}
clutches_of_deads1=dead_wRing_knownDd_blood%>%
  select(BornClutchId)%>%
    distinct(BornClutchId)


all_sibs=bird%>%
  right_join(clutches_of_deads1)%>% #645 
  anti_join(sequenced_ids, by = "RingId")%>%
  anti_join(dead_wRing_knownDd_blood_IDs, by = 'RingId')

ringed_sibs=
  all_sibs%>%
  left_join(lab)%>%
  filter(SampleType%in%c("blood","Blood cell"))

n_distinct(ringed_sibs$RingId)

ringed_sibs
```

## STAGE 1.3 ##

Unringed siblings from proposed clutches
(so ids that are unringed from the clutches that had an individual die in)
70 ids 

```{r}

unringed_sibs=bird%>%
  right_join(clutches_of_deads1)%>% 
  filter(RingId=="")
  
unringed_sibs[1:7, c(1:4,6,8)]
```





Ids with high ibcs where we could seq siblings too 

```{r}

##Fgrm and FROH from elo
inb=read.table("Inbreeding_depression_owls/All3085_FuniWE_FHBD512g.txt", header = T)%>%
  rename(RingId=INDVs)

top_ibcs=inb%>%
  filter(FuniWE > quantile(FuniWE, 0.9))%>% # top 90% ibcs
  left_join(bird)%>%
  filter(as.Date(HatchDate, format = "%d/%m/%Y") > as.Date("2010-01-01")) # hatched after / in 2010

n_distinct(top_ibcs$BornClutchId) ##  different clutches that have highest ibcs 

```

Siblings of (previously seq) individuals with high inbreeding coefficients 

```{r}
inbred_clutches=top_ibcs%>%
  select(BornClutchId)%>%
  distinct(BornClutchId)%>%
  na.omit()



inbred_sibs=bird%>%
  right_join(inbred_clutches)%>% #
  anti_join(sequenced_ids, by = "RingId")%>%
  anti_join(dead_wRing_knownDd_blood_IDs, by = 'RingId')%>%
  anti_join(ringed_sibs, by = 'RingId')%>%
  left_join(lab,  by = "RingId")%>%
  filter(SampleType%in%c("blood","Blood cell"))

n_distinct(inbred_sibs$RingId) #
n_distinct(inbred_sibs$BornClutchId) #


```

Unringed siblings of (previously seq) individuals with high inbreeding coefficients 
(unsure of type of sample)


```{r}

inbred_sibs_noring=bird%>%
  right_join(inbred_clutches)%>% #
  filter(RingId=="")

n_distinct(inbred_sibs_noring$BirdId) #
n_distinct(inbred_sibs_noring$BornClutchId) #


```

All individuals born after 2010 with no RingId

(PROBLEM: not sure how to link bird id to lab info)
```{r}
All_after2010_noRing=bird%>%
    filter(as.Date(HatchDate, format = "%d/%m/%Y") > as.Date("2010-01-01"))%>% # hatched after / in 2010
  filter(RingId=="")
 
n_distinct(All_after2010_noRing$BirdId)


```

Of those, those with known death date:

```{r}
dead_noRing_knownDd=All_after2010_noRing%>%
  filter(!(DeathDate==""))

n_distinct(dead_noRing_knownDd$BirdId)

```

So 400 ids with no known death date but no ringID born after 2010 - so perhaphs they disapeared??

need to find ids where last observation was before the siblings and within ~40 days of hatching

```{r}
missing_noRing=All_after2010_noRing%>%
  filter(DeathDate=="")%>%
  filter(as.Date(LatestBirdObservationDate, format = "%d/%m/%Y") < as.Date(HatchDate, format = "%d/%m/%Y")+20) 


#n_distinct(dead_noRing_knownDd$BirdId)

```



```{r}

all_missingIDs_unknownDd <- bird %>%
  filter(!is.na(BornClutchId)) %>%
  filter(as.Date(HatchDate, format = "%d/%m/%Y") > as.Date("2010-01-01")) %>% # hatched after / in 2010
  filter(DeathDate == "" | is.na(DeathDate)) %>%
  group_by(BornClutchId) %>%
  mutate(
    LatestBirdObservationDate = as.Date(LatestBirdObservationDate, format = "%d/%m/%Y"),
    hatchdate = as.Date(HatchDate, format = "%d/%m/%Y") ) %>%
  mutate(
    max_sibling_obs = max(LatestBirdObservationDate[LatestBirdObservationDate <= hatchdate + 60]) ) %>%
  filter(
    LatestBirdObservationDate < max_sibling_obs &
    LatestBirdObservationDate < hatchdate + 30) %>%
  ungroup()



```






Also ids that went missing but do still have a ringID?
(or maybe these ids have estimated death days)
#############################################################################

CONSIDERATIONS:

- remove cross fostered ids? (how many were CF after 2010 anyway?)
- check parents of clutches whether theyve also been sequenced 
